Stack Safety as a Security Property

### 2020-07-09 main.tex
Revamp of 6

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\mpstatename \in \MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with respect to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

@@

Both the integrity and confidentiality components of local state
encapsulation concern themselves only with the accessibility of data, and
may still hold on a wide range of programs that violate expected control
flow. Say, if {\tt g} managed to jump in the middle of the
code of {\tt h} with its call, confidentiality and integrity would
still be satisfied.
%
However, that is still unwanted behavior: for example, a caller (or a
callee) should not be able to bypass a password check by jumping (or
returning) in the middle of a block of code --- even if such an execution
would ostensibly satisfy local state encapsulation.
%
Informal definitions of stack safety usually \apt{evidence?} aim to prevent such unwanted
flows: every jump between functions should either be a call or a
return, and each return should be to the instruction after its
corresponding call.

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (call map $\callmap$, function map $\codemap$, return map $\retmap$, and entry
map $\entmap$): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state $\mpstatename$ and any adjacent pair
of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
trace from $\mpstatename$ whose current instruction
belongs to different functions according to the function map,
$\mach_1$ is either a valid call or a valid return:

@@

For example, consider changing line 5 of the running example to a $\jalr$,
$\PCname \leftarrow r_4$, with effect $\PCname \leftarrow 42$, causing {\tt main}
to jump into the middle of {\tt h}. 
This violates control separation because the instruction that causes a
function switch is not a valid call (or a return) according to the
call map.

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state $\mpstatename$
and any n-transition \((\mach_c,\pol_c)\) and
\((\mach_e, \pol_e)\) in the induced trace from
$\mpstatename$, $\mach_e$ is a member of the entry map:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

The changed program described just above also violates entry integrity,
because the target (instruction 42) of the new $\jalr$ is not an established entry point
according to the entry map.

@@

\newcommand*{\Unmatched}{\mathit{um}}

Return integrity captures the expected behavior of marked returns with
respect to their call points. Intuitively, it says that the marked return
matching a call $\mach_c$ should satisfy $\ret{\mach_c}$.
Observe that, given an MP-trace $\mpstate{\mach_c}{\pol}\MPT$, if $\mach_c$ is a call state then
the matching return will be the first \emph{unmatched} return in $\MPT$. So
we begin by defining an auxiliary ``unmatched'' relation between two MP-traces.

\definition
We write $\Unmatched~\MPT~\MPT'$ if $\MPT'$ is
obtained from $\MPT$ by dropping elements until we reach the first unmatched return
(according to the return map).  Formally, $\Unmatched$ is defined coinductively as follows:
\begin{itemize}
\item A trace that starts with a return is related to itself:

\begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}
\end{minipage}

@@

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))
              \hspace{.5cm}\Unmatched ~ \MPT ~ \MPT'\)}
           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}

\item Finally, if a trace starts with a call, then we must discard the first
  unmatched return from its tail to get to the real one:

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
\apt{Could cut formal definition for space.}

\definition%{Return Integrity}
A system enjoys {\em return integrity}
% with respect to a call map \(\callmap\) and return map \(\retmap\)
if, for any initial state pair 
$\mpstatename$ and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow \MPT\) and \(\mach_r\)
is the first unmatched return in $\MPT$, then $\mach_r$ corresponds
to a return state for $\mach_c$:

\[
\exists p_r.\Unmatched ~ \MPT ~ (m_r,p_r)\MPT' \Rightarrow \ret{\mach_c}\ \mach_r.
\]
\apt{I've tried to tidy this, but I don't think it corresponds to the Coq.
  Shouldn't we apply $\Unmatched$ to the tail of $\MPT$?
  And what about the no-unmatched-return-at-top-level clause?}

For example, if we modified the running example by adding
an instruction $ \rra \rightarrow \rra - 19$
between 45 and 46, this would cause the return from {\tt h} to go to instruction 8 and
return to {\tt main} instead of {\tt g}. While the instruction that
causes the return would still be a valid return (satisfying control
separation), the state after the return would not correspond to an
unmatched return for {\tt g}'s call to {\tt h}, violating return
integrity.

@@

The above properties are somewhat orthogonal, but together they
capture a standard notion of control flow for a stack: functions keep
their code separate outside of calls and returns, calls respect
explicit entry points, and returns respect implicit entry points
corresponding to their call.  A system that enjoys all three
properties (with respect to a function map, call map, return map,
and entry map) is said to enjoy
{\em well-bracketed control flow} (with respect to those maps).

@@

\paragraph*{Strong Stack Safety}
%
This conservative policy does not only satisfy the stack safety above,
but also a provably stronger one, which at each step of the
computation performs integrity and confidentiality tests for a stack
of running variants, which are kept in lockstep with the original
machine.

@@

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step,
  denoted $\vm{C} \stepstoobs{\obs} \mach'$, while
  satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
  and stack confidentiality test $\testConf{\mach}{\mach'}$
  with respect to the original machine state. We write
  this $\testVS{(\mach, \pol)}{\vs}$.

\definition
%  Let $\callmap$ a call map and $\codemap$ a function map.
  Given a state $(\mach, \pol) \in \MPS$ that is well-formed w.r.t. a
  variant stack $\vs$, we say that an MP-trace induced from $(\mach,
  \pol)$ satisfies the \emph{strong stack safety} at a given call
  depth as follows.
% Refactor repetitive notation before case split
  In all cases that let $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$.
  %and assume (unless the step is a return) that the
  %variant stack can step, $\vs \hookrightarrow \vs'$, i.e., all its variant
  %machines can step.

%  Let $\callmap$ a call map and $\codemap$ a code map. Given a state $(\mach,
%  \pol) \in \MPS$ that is well-formed w.r.t. a variant stack $\vs$ and induces
%  an MP-trace $(\mach, \pol) \hookrightarrow \MPT$, we say that the \emph{stack
%  testing property} is satisfied at call depth $d$, and write
%  $\mathit{stack}_T ~ d ~ \MPT ~ \vs$, as follows:

@@

  \item
%
    If the machine takes a \emph{call step} and $\contour$ is the contour at
    the call then,
%
    (1) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
    satisfied,
%
    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
    is satisfied with regard to an arbitrary $\contour$-variant
    of $\mach$,
%
    and (4) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{(\vm{\contour} : \vs')}$ is coinductively satisfied, for an arbitrary
    $\contour$-variant of the machine,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

  \item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\testProp{d}{[(\mach, \pol)]}{\vs}$, for any $d$ up to the size of
    $\vs$. (That is, a violation was detected $d$ levels into the stack.
    Elements are created and pushed to the variant stack on calls and popped
    from it on returns, so its size always reflects the depth of the program
    stack.)

  \item If the top variant of the stack (and only the top variant) returns,
    %i.e., $\mathit{ret}(\head(\vs), \mach)$ holds,
    the property is safisfied for
    the singleton trace $[(\mach, \pol)]$ at depth equal to the size of the
    stack: $\testProp{|\vs|}{[(\mach, \pol)]}{\vs}$. \rb{Explain the
    use of this vs the return case below.}

  \item If
%
    (1) the machine takes a \emph{non-call, non-return step},
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied, and
%
    (3) the stack safety test $\testProp{d}{((\mach', \pol') : \MPT)}{\vs'}$
    is coinductively satisfied % at depth $d$
    %% and for a trace starting continuing after the step, $\head(\MPT) = (\mach',
    %% \pol')$,
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach', \pol') :
    \MPT)}{vs}$.
    %% \rb{omitting stack size condition}

    %% the contour $\contour$ is updated based on the top of the current machine.
    %% For the property to hold for a trace $(\mach, \pol) \cdot \MPT$ at depth $d$
    %% the eager integrity test must hold for $\contour$ at the call,
    %% the eager confidentiality test must hold for all variants at the call,
    %% including the new variant that is added to the top of the stack,
    %% and the property recurses on an arbitrary variant at the new contour
    %% at depth $d$ with trace $\MPT$.

  \item
%
    If (1) the machine takes a \emph{return step}, so the top of the variant
    stack (and only the top variant) returns,
%
    (2) the variant stack test $\testVS{(\mach, \pol)}{\vs}$ is
    satisfied,
%
    (3) the stack safety test $\testProp{d}{((\mach', \pol') :
    \MPT)}{\mathit{tail}(\vs)}$ is coinductively satisfied
%
    {\em then} we conclude $\testProp{d}{((\mach, \pol) : (\mach',
    \pol') : \MPT)}{vs}$.

  \item If the machine takes a \emph{simultaneous return-call step}, that is,
    the instruction to which execution returns immediately after completing a
    call is itself a call instruction, the \emph{call step} and \emph{return
    step} cases are combined.

@@

\definition
%
  A trace $\MPT$ satisfies our \emph{top-level stack testing property} if it
  satisfies the eager testing property at depth 1 with respect to the initial
  variant, $\testProp{1}{\MPT}{\vm{\contour_0}}$.
%
  \rb{I don't think the initial variant has been defined, but it's
    straightforward.}

By using property-based testing techniques, we can determine with a high degree
of confidence that an implementation of the conservative micropolicy satisfies
this eager testing property. Although we have not proved that the conservative
policy guarantees the stack safety property, the testing property is provably
stronger than the stack safety property \rb{theorem?} and, by this implication,
we can conclude that the conservative policy also enforces stack safety.

% The properties we give are indeed enforceable.

\rb{What is the strongest property enforced by the conservative policy?}

\paragraph*{End-to-end stack safety property}

Integrity

Confidentiality as a form of noninterference

\paragraph*{Eager vs lazy testing}

The conservative policy we just described, while testable and enforceable, is
too slow to be of practical use. What we want, instead, is a somewhat more
permissive policy that moreover can be efficiently implemented. Most of the
performance overhead incurred by this kind of \emph{eager policy} steps from the
need to set stack activation tags as a frame is created and reset them as it is
destructed, as \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
mitigate those costs, they propose optimizations like Lazy Tagging and Lazy
Clearing that relax and defer some of the checks of an eager policy, being more
permissive in their allowed operations and catching violations if and when their
effect becomes manifest.
\rb{How much detail to go into? Refer to examples if we have the space?}

These \emph{lazy policies} admit more efficient implementations while being
harder to test, given that potential violations may or may not materialize,
requiring in turn additional bookkeeping. However, a critical observation is
that the depth-based tagging scheme discussed so far with fail when combined
with a lazy policy. The reason for this is that because tags are not set and
cleared on function entry and exit points, stale tags may persist after a given
call, and a subsequent call sharing the same depth as the old call may access
memory contents tagged with the stale tags of a previous activation.
\rb{Again, example?}

@@

      A system enjoys {\em observable stack integrity} with respect to
      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with respect to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with respect to a call map
      and a function map if it enjoys both observable stack integrity and
      observable stack confidentiality with respect to them.

>>>>

We write $\MPS$ for the set of all pairs of machine states and policy
states.
%
\[\MPS = \MACHS \times \POLS\]
%
We lift the policy step function to operate on such pairs
by combining it with the regular step function for machine states.

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
We say that a
system satisfies {\em stack confidentiality}
% with respect to $\callmap$ and $\codemap$
if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $\MPT$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ 
in $\MPT$, the following holds. Let $\MPT_{\mathit{call}}$ be the prefix where
\(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_{\mathit{call}} | \ret{\mach_c}\).
For each variant state \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)}
\mach_e\), let \(\machT\) be the similarly prefixed trace such that
\(\mach_e' \hookrightarrow \machT\). Then trace confidentiality holds on
\(\MPT_{\mathit{call}}\) and \(\machT\) with regard to \(\mathit{Cof}(\mach_c, n)\)
and \(R = \ret{\mach_c}\):

@@

Both the integrity and confidentiality components of local state
encapsulation concern themselves with the accessibility of data, and
may hold on a wide range of programs that violate expected control
flow. Say, if {\tt g} managed to jump in the middle of the
code of {\tt h} with its call, confidentiality and integrity would
still be satisfied.
%
However, that is still unwanted behavior: for example, a caller (or a
callee) should not be able to bypass a password check by jumping (or
returning) in the middle of a block of code --- even if such an execution
would ostensibly satisfy local state encapsulation.
%
Stack safety is traditionally associated with preventing such unwanted
flows: every jump between functions should either be a call or a
return, and each return should be to the instruction after its
corresponding call.

To formally capture this aspect of stack safety, we introduce three
security properties that rely on the underlying knowledge of the
program structure (call maps, function maps, return maps, and entry
maps): {\em control separation} ensures the validity of transition
sources between functions, {\em entry integrity} ensures the validity
of call targets, and {\em return integrity} ensures the validity of
return targets.

\definition{Control Separation}

A system enjoys {\em control separation} with regard to a function map
\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if for any initial state \((\mach_0, \pol_0)\) and any adjacent pair
of states \((\mach_1,\pol_1)\) and \((\mach_2,\pol_2)\) in the induced
trace from \(\mpstate{\mach_0}{\pol_0})\) whose current instruction
belongs to different functions according to the function map,
either $\mach_1$ is a valid call or a valid return:

@@

For example, consider the unsafe substitution:
\[
5 : ~ \jalr ~ r_4 ~ \rra ~ 0
\]
This violates control separation as the instruction that causes a
function switch is not a valid call (or a return) according to the
call map.

\definition{Entry Integrity}

A system enjoys {\em entry integrity} with regard to a call map
\(\callmap\) and entry map \(\entmap\) if for any initial state
\((\mach_0, \pol_0)\) and any n-transition \((\mach_c,\pol_c)\) and
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a member of the entry map:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

The same unsafe substitution violates entry integrity
\[
5 : ~ \jalr ~ r_4 ~ \rra ~ 0
\]
as its target (instruction 42) is not an established entry point for
{\tt h} according to the entry map.

@@

\definition{Return Integrity}

\newcommand*{\Unmatched}{\mathit{Unmatched}}

Return integrity captures the expected behavior of marked returns with
regard to their call points. Given a call map $\callmap$ and a return
map $\retmap$, we begin by defining a coinductive relation
$\Unmatched$ over two $\MPOTRACE$ $\MPT$ and $\MPT'$, if $\MPT'$ is a
suffix of $\MPT$ whose head is the first unmatched return in $\MPT$.

Formally:
\begin{itemize}
\item A trace that starts with a return is related to itself:

  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol} ~ \mpstate{\mach}{\pol}\)}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
              {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \mpstate{\mach}{\pol}\MPT\)}

@@

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))
              \Unmatched ~ \MPT ~ \MPT'\)}
           {\(\Unmatched ~ \mpstate{\mach}{\pol}\MPT ~ \MPT'\)}

\item Finally, if a trace starts with the call, then we must discard the first
  unmatched return to get to the real one:

  \judgmentbr{\(\callmap(\mach(\PCname)) = n \hspace{.5cm} \neg \retmap(\mach(\PCname)) \)}
             {\(\MPT = (m,p)\MPT_{\mathit{call}}\) \quad \(\Unmatched ~ \MPT_{\mathit{call}} ~ \MPT_{\mathit{matched}}\)}
             {\(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\) \quad
              \(\Unmatched ~ \MPT_{\mathit{tail}} ~ \MPT_{\mathit{unmatched}}\)}
            {\(\Unmatched ~ \MPT ~ \MPT_{\mathit{unmatched}}\)}
\end{itemize}
A system enjoys {\em return integrity} with regard to a call map
\(\callmap\) and return map \(\retmap\) if for any initial state pair
\((\mach_0, \pol_0)\) and any state pair \((\mach_c, \pol_c)\)
corresponding to a call in its induced trace, if
\(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT\) and \(\mach_r\)
is the first unmatched return such that for some $\pol_r$
\[ \Unmatched ~ \MPT ~ (m_r,p_r)\MPT',\]
then $\mach_r$ corresponds to a return state for $\mach_c$:
\[\ret{\mach_c}\ \mach_r\]

For example, if we added an instruction between 45 and 46:
\[
45.5: ~ \addi ~ \rra ~ r_0 ~ \negate 19
\]
This would cause the return from {\tt h} to go to instruction 8 and
return to {\tt main} instead of {\tt g}. While the instruction that
causes the return would still be a valid return (satisfying control
separation), the state after the return would not correspond to an
$\Unmatched$ return for {\tt g}'s call to {\tt h}, violating return
integrity.

@@

The above properties are somewhat orthogonal, but together they
capture a standard notion of control flow for a stack: functions keep
their code separate outside of calls and returns, calls respect
explicit entry points, and returns respect implicit entry points
corresponding to their call.  A system that enjoys them with regard to
a code, a call, a return, and an entry map is said to enjoy
{\em well-bracketed control flow} with regard to those maps.

@@

\paragraph*{Strong Stack Safety}

\leo{Maybe lockstep?}

This conservative policy does not only satisfy the stack safety above,
but also a provably (in Coq) stronger one that enforces local state
encapsulation throughout every calls execution. To encode that
property, we not only keep track of the execution of one machine
(whose trace we're trying to characterize as safe), but of an entire
stack of machine variations corresponding to nested calls of the
original machine. For each step of the original machine, we take a
step in every variant machine (thus, keeping them in lockstep
execution) and enforce a single-step variant of integrity and
confidentiality tests. In addition, every time a call is made we push
a variation of the original machine state at the time of that call on
top of the variant stack (just like the stack confidentiality property
enforced trace confidentiality for every transition), and every time
the machine returns from a call we pop the top variation (to account
for stack confidentiality taking the prefix of the transition trace
until its return).

We start by defining the integrity and confidentiality tests, as well
as the variant stack, before giving the full definition of strong
stack safety.

@@

We also need two auxilliary notations that relate to variant stacks.
% 
First, strong safety will check that the basic tests above hold for every
machine in the variant stack.  Given a machine that steps $(\mach,
\pol) \hookrightarrow_P (\mach', \pol', \obs)$, say that a variant
stack $\vs$ safisfies the \emph{variant stack test} if for each
variant element $\vm{C}$ (which is a variation of $\mach$ for $C$),
$\vm{C}$ can take a step, denoted $\vm{C} \stepstoobs{\obs} \mach'$,
while satisfying the stack integrity test $\testInt{C}{\mach}{\mach'}$
and stack confidentiality test $\testConf{\mach}{\mach'}$ with regard
to the original machine state. We will write $\testVS{(\mach,
  \pol)}{\vs}$ for this compound check.
%
Second, we will write $\vs \stepsto \vs'$ if every variant machine
in $\vs$ steps to the corresponding machine in $\vs'$.
%

We are now ready to define strong stack safety coinductively.

\definition Given a state $(\mach, \pol) \in \MPS$ and a variant stack
$\vs$ in which every element $\vm{\contour}$ is a variation of $\mach$
with respect to its contour, we say that an MP-trace induced from
$(\mach, \pol)$ satisfies the \emph{trace strong stack safety}, and write
$\testProp{}{\MPT}{\vs}$, if:

@@

\item
  If the machine takes a \emph{non-call, non-return step}
  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)\MPT,\]
  then the variant stack test must be satisfied:
  \[\testVS{(\mach, \pol)}{\vs},\]
  and trace strong stack safety must be coinductively satisfied for
  the remainder of the trace and the variant stack $vs'$ that
  steps to ($\vs \stepsto \vs'$):
  \[ \testProp{}{((\mach',\pol',\obs)\MPT)}{\vs'}\]

\item
  If the machine takes a \emph{call step} and $\contour$ is the contour at
  the call
  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \callmap(\mach) = n,\]
%
  then the variant stack test $\testVS{(\mach, \pol)}{\vs}$ must be
  satisfied:
  \[\testVS{(\mach, \pol)}{\vs},\]
%
%    (2) the stack integrity test $\testInt{\contour}{\mach}{\mach'}$ is
%    satisfied and the stack confidentiality test $\testConf{\mach}{\vm{\contour}}$
%    is satisfied with regard to an arbitrary $\contour$-variant
%    of $\mach$,
%
  and trace strong stack safety must be coinductively satisfied
  for the remainder of the trace and for every variant stack
  that is the result of adding a variation of $\mach'$ with
  respect to $C$ at the top of the stack $\vs'$ that $\vs$ steps
  to:
  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\vm{\contour} : \vs')}\]

\item
%
  If the machine $\mach$ takes a \emph{return step},
  \[(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs) \wedge \retmap(\mach),\]
  %
  the top of the variant stack also returns,
  the tail of $\vs$ satisfies the stack safety test
  \[\testVS{(\mach, \pol)}{\mathit{tail}(\vs)},\]
  %
  and trace strong stack safety must be coinductively satisfied for
  the remainder of the trace and for the tail of the varant stack
  that $\vs$ steps to:
  \[\testProp{}{((\mach', \pol', \obs)\MPT)}{(\mathit{tail}(\vs'))}\]

\item If the machine \emph{fail-stops}, i.e., $(\mach, \pol) \hookrightarrow_P
    (\mach, \pol, \tau)$, the property is safisfied for the singleton trace:
    $\testProp{}{[(\mach, \pol)]}{\vs}$.

%  \item If the machine takes a \emph{simultaneous return-call step}, that is,
%    the instruction to which execution returns immediately after completing a
%    call is itself a call instruction, the \emph{call step} and \emph{return
%    step} cases are combined.

@@

\definition 
%
A system satisfies {\em strong stack safety} if for every initial machine $\mach$,
the trace induced from $(\mach, \pinit{\mach})$ satisfies trace strong stack safety with
respect to the empty stack.

\theorem
Strong stack safety implies stack safety.

\proof
By coinduction on trace strong stack safety.
\leo{3000 lines Coq?}
\leo{I can write a TON about this proof, but I'm not sure how much is interesting/how much space we have}.

\paragraph*{Eager vs Lazy Stack Safety}

The conservative policy we just described, while testable and enforceable, is
too slow to be of practical use. What we want, instead, is a somewhat more
permissive policy that moreover can be efficiently implemented. Most of the
performance overhead incurred by this kind of \emph{eager policy} steps from the
need to set stack activation tags as a frame is created and reset them as it is
destructed, as \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
mitigate those costs, they propose optimizations like Lazy Tagging and Lazy
Clearing that relax and defer some of the checks of an eager policy, being more
permissive in their allowed operations and catching violations if and when their
effect becomes manifest.

\leo{Not sure how much of the following could go after the lazy stack safety property}
These \emph{lazy policies} admit more efficient implementations while being
harder to test, given that potential violations may or may not materialize,
requiring in turn additional bookkeeping. However, a critical observation is
that the depth-based tagging scheme discussed so far with fail when combined
with a lazy policy. The reason for this is that because tags are not set and
cleared on function entry and exit points, stale tags may persist after a given
call, and a subsequent call sharing the same depth as the old call may access
memory contents tagged with the stale tags of a previous activation.
\rb{Again, example?}

@@

      A system enjoys {\em observable stack integrity} with regard to
      \(\callmap\) and \(\codemap\) if for any initial state \(\mach_0\) and
      each transition with entry state \(\mpostate{\mach_e}{\pol_e}{\obs_e}\)
      and contour \(\contour\) in the induced trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsInt}_\contour ~ \MPT\).

@@

      From observable trace confidentiality we construct {\em observable stack
      confidentiality}. A system enjoys {\em observable stack confidentiality}
      with regard to \(\callmap\) and \(\codemap\) if for any initial state
      \(\mach_0\) and each transition with entry state
      \(\mpostate{\mach_e}{\pol_e}{\obs_e}\) and contour \(\contour\) in the
      induced trace from \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), if
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_e}\),
      then \(\text{\sc ObsConf}_\contour ~ \MPT ~ \machT\).

    \definition{Observable Stack Safety}
      A system enjoys {\em observable stack safety} with regard to a call map
      and a function map if it enjoys both observable stack integrity and
      observable stack confidentiality with regard to them.
